Effect of lakritsi on agressiveness: a Moomin perspective
=========================================================



Context of the study
--------------------



Haribo *et al.* have suggested a correlation between the 
consumption of **liquorice** and **agressive behaviour**, but 
their previous study on Smurfs (*Schtroumpfus peyoii*) remained inconclusive.

Haribo *et al.* study had 3 major flaws:
- they did not describe how they measured agressiveness
- most Smurfs escaped before the end of the experiment
- Smurfs have a low expression level of liquorice receptors, as 
was later evidenced:
    1. in the brain
    2. in the liver
    3. in the muscle



Experimental design
-------------------



We decided to test their hypothesis using a dataset collected on Moomins and
other related characters. Our approach includes:
- male and female individuals from different species
- different levels of liquorice consumption among individuals
- paintball activity as a proxy for agressiveness



Individuals used in this study
------------------------------



```{r loadMoominData}
# Load Moomin data
moomin = read.table("../data/moomin_data", header = T, sep ="\t")
moomin
```



Results
-------



### Paintball activity



For each individual, the number of balls used during a 180min 
paintball game was monitored.

```{r loadPaintballData}
# Load paintball data
paintball = read.table("../data/paintball_data", header = T, 
                       sep ="\t")
head(paintball)
```



### Example of activity: Sniff (Nipsu)



This plot shows the amount of balls used by Sniff during the
game.

```{r plotSniffActivity}
# Extract Sniff activity
sniff_activity = subset(paintball, paintball$user == "Sniff")
plot(sniff_activity$time.min, sniff_activity$balls.used, type = "l")
```

The next plot uses `ggplot2` for fancier graphics.

```{r plotSniffActivity.ggplot2}
library(ggplot2)
ggplot(sniff_activity, aes(x = time.min, y = balls.used)) +
   geom_line()
```



### Activity of all individuals



The next plot uses the data for all `r nrow(paintball)` 
records in the paintball tables.

```{r plotActivityAll, fig.width = 8, fig.height = 6}
ggplot(paintball, aes(x = time.min, y = balls.used, col = user)) +
   geom_line()
```



### Effect of lakritsi on paintball activity

We have an estimation of the average lakritsi consumption per
day for each individual. Consumption ranges from 
`r min(moomin$liquorice.kg.day)` to 
`r max(moomin$liquorice.kg.day)` kg per day.



```{r mergeTables, echo = FALSE}
all_data = merge(moomin, paintball, by.x = "english", 
                 by.y = "user")
```



```{r calculateTotalNumberBalls, echo = FALSE}
total.n.balls = vector()
for (individual in moomin$english) {
  total = max(subset(paintball, 
                     paintball$user == individual)$balls.used)
  total.n.balls = c(total.n.balls, total)
}
summary_moomin = data.frame(moomin$english,
                            moomin$liquorice.kg.day,
                            total.n.balls)
names(summary_moomin) = c("individual", "liquorice.kg.day",
                          "balls.game")
```



The next plot uses color-coding to depict the average consumption
of lakritsi in kg/day.

```{r plotActivity.col.lakritsi, fig.width = 8, fig.height = 5}
ggplot(all_data, aes(x = time.min, y = balls.used, group = english)) +
   geom_line(aes(col = liquorice.kg.day))
```



We can also look at the total number of balls used in relation to the
liquorice consumption.

```{r plotTotalBallsLiquorice, fig.width = 8, fig.height = 5}
ggplot(summary_moomin, aes(x = liquorice.kg.day, y = balls.game,
                           col = individual))+ geom_point()
```



The Spearman's rho test gives:

```{r SpearmanTest}
cor.test(summary_moomin$liquorice.kg.day, summary_moomin$balls.game,
         method = "spearman")
```



Comparison of the studies
-------------------------



Haribo *et al.* | our study
-----|-----
Smurfs | Moomins
unknown proxy | paintball activity
potential conflict of interest | no conflict of interest
